The imprint of global magma oceans on exoplanet demographics

### 2023-01-04 src/tex/ms.tex
minor edits

\begin{note}
    The power from the host star per unit area at the position of a planet or instellation $S$ in units of Earth's insolation is given by
    \begin{equation}
        \frac{S}{S_\oplus} = \left(\frac{L_\star}{L_\odot}\right) \left(\frac{au}{a}\right)^2 .
    \end{equation}
    Here, we assume global redistribution of incoming flux and a fixed albedo of $0.3$, comparable to Earth's [CITE].
    The absolute value of $S_\oplus$, by which all synthetic planets are scaled, is then \SI{238}{\watt\per\square\meter}.
%    We further define the solar-equivalent semi-major axis $a_{eff} = a (L_\star/L_\odot)^{-1/2}$, at which a planet experiences the same instellation as a Solar System planet at an orbital distance~$a$.
\end{note}

@@

\subsection{Survey simulations}
\begin{note}
The survey module of Bioverse converts the synthetic planet sample into a set of uncertainty-laden measurements of some parameters of a subset of that sample.
   This task includes selection of the targets, application of detection biases, and conducting simulated measurements.
   All of the above are specific to the particular survey.

@@

As our procedure involves random sampling of the model parameters $\theta$, we need to define the probability of obtaining a dataset given the model parameters, i.e., a likelihood function.
We assumed here that the individual moving averages $SMA_i$ are measured with a normally distributed uncertainty $\sigma_{SMA_i}$ and adopted a normal distribution
\begin{equation}
    \mathcal{L}(SMA \mid \boldsymbol{\theta})=\prod_{i}^{N} \frac{1}{\sqrt{2 \pi \sigma_{SMA_i}^{2}}} \exp \left(-\frac{\left(SMA_i - h\left(\boldsymbol{\theta}, S_i\right)\right)^{2}}{2 \sigma_{SMA_i}^{2}}\right).
\end{equation}
Here, $h\left(\boldsymbol{\theta}, S_i\right)$ corresponds to the functional form of the runaway greenhouse hypothesis~(Eqn.~\ref{eq:rgh_hypo}).
We tested this hypothesis against the null hypothesis $h_\mathrm{null} (\theta, S) = \theta$, which states that there is no relationship between the measured planet radii and received instellation.
\end{note}

\subsubsection{Bayesian model comparison}
We are now ready to assess the relative plausibility of $H_{\mathrm{rgh}}$ and $H_0$ given the synthetic data we have generated, assigning equal a~priori~ probabilities to these models.
This is done by comparing the Bayesian evidence $\mathcal{Z}$ of the models, which we estimate with the nested sampling~\citep{Skilling2004} algorithm \emph{dynesty}~\citep{Speagle2020}.
Our criterion to reject the null hypothesis is
\begin{equation}
\Delta \ln \mathcal{Z}  = \ln \mathcal{Z} - \ln \mathcal{Z}_\mathrm{null}  > 3.
\end{equation}
The chosen threshold roughly corresponds to the $p = 0.05$ threshold for statistical significance popular in frequentist approaches.\todo{CITATION}

@@

\section{Discussion}\label{sec:discussion}
\todo[inline]{\textbf{Potential discussion points: } what is the fraction of Kepler planets w/ runaway GH.}

@@

\subsection{Detectability of the runaway greenhouse threshold}
\todo[inline]{the demographic imprint is detectable. Under what conditions?}

@@

The key findings of our work are as follows:
\begin{enumerate}
    \item The predicted tipping point in planetary climates causes a discontinuity in the radius and density distribution of small (DEFINITION) exoplanets.
    \item Its demographic imprint of the runaway greenhouse threshold should be detectable with transit measurements of a large enough sample and sufficient precision.
    \item We find that the planned PLATO transit survey will provide a sufficient sample to test the runaway greenhouse hypothesis.
    \item The diagnostic power of transit missions in testing this hypothesis can be maximized by ...
    \item A detection of the feature will constrain the water inventory of rocky exoplanets, providing important context in the assessment of their habitability.
\end{enumerate}

>>>>

\todo[inline]{ARNAUD: revise definition of $S$ following our discussion on 2022-12-22. Introduce net instellation and clarify what is meant by it.}
\begin{note}
    The power from the host star per unit area at the position of a planet or instellation $S$ in units of Earth's insolation is given by
    \begin{equation}
        \frac{S}{S_\oplus} = \left(\frac{L_\star}{L_\odot}\right) \left(\frac{au}{a}\right)^2 .
    \end{equation}
    Here, we assume global redistribution of incoming flux and a fixed albedo of $0.3$, comparable to Earth's [CITE].
    The absolute value of $S_\oplus$, by which all synthetic planets are scaled, is then \SI{238}{\watt\per\square\meter}.
%    We further define the solar-equivalent semi-major axis $a_{eff} = a (L_\star/L_\odot)^{-1/2}$, at which a planet experiences the same instellation as a Solar System planet at an orbital distance~$a$.
\end{note}

@@

\subsection{Survey simulations}
\begin{note}
The survey module of Bioverse converts the synthetic planet sample into a set of uncertainty-laden measurements on a subset of that sample.
   This task includes selection of the targets, application of detection biases, and conducting simulated measurements.
   All of the above are specific to the particular survey.

@@

As our procedure involves random sampling of the model parameters $\theta$, we need to define the probability of obtaining a dataset given the model parameters, i.e., a likelihood function.
We assumed here that the individual moving averages $SMA_i$ are measured with a normally distributed uncertainty $\sigma_{SMA_i}$ and adopted a normal distribution
\begin{equation}
    \mathcal{L}(SMA \mid \boldsymbol{\theta})=\prod_{i}^{N} \frac{1}{\sqrt{2 \pi \sigma_{SMA_i}^{2}}} \exp \left(-\frac{\left(SMA_i - h\left(\boldsymbol{\theta}, S_i\right)\right)^{2}}{2 \sigma_{SMA_i}^{2}}\right).
\end{equation}
Here, $h\left(\boldsymbol{\theta}, S_i\right)$ corresponds to the functional form of the runaway greenhouse hypothesis~(Eqn.~\ref{eq:rgh_hypo}).
We tested this hypothesis against the null hypothesis $H_0 (\theta, S) = \theta$, which states that there is no relationship between the measured planet radii and received instellation.
\end{note}

\subsubsection{Bayesian model comparison}
We are now ready to assess the relative plausibility of $H_{\mathrm{rgh}}$ and $H_0$ given the synthetic data we have generated, assigning equal a~priori~ probabilities to these models.
This is done by comparing the Bayesian evidence $\mathcal{Z}$ of the models, which we estimate with the nested sampling~\citep{Skilling2004} algorithm \emph{dynesty}~\citep{Speagle2020}.
Our criterion to reject the null hypothesis is
\begin{equation}
\Delta \ln \mathcal{Z}  = \ln \mathcal{Z}_\mathrm{rgh} - \ln \mathcal{Z}_0  > 3.
\end{equation}
The chosen threshold roughly corresponds to the $p = 0.05$ threshold for statistical significance popular in frequentist approaches.\todo{CITATION}

@@

\section{Discussion}\label{sec:discussion}
\todo[inline]{\textbf{Potential discussion points: } what is the fraction of \kepler\ planets w/ runaway GH.}

@@

\subsection{Detectability of the runaway greenhouse threshold}
\todo[inline]{the demographic imprint is detectable. Under what conditions?\\
Mention multiplanet systems.}

@@

The key findings of our work are as follows:
\begin{enumerate}
    \item The predicted tipping point in planetary climates causes a discontinuity in the radius and density distribution of small (DEFINITION) exoplanets.
    \item The demographic imprint of the runaway greenhouse threshold should be detectable with transit measurements of a large enough sample and sufficient precision.
    \item We find that the planned PLATO transit survey will provide a sufficient sample to test the runaway greenhouse hypothesis.
    \item The diagnostic power of transit missions in testing this hypothesis can be maximized by ...
    \item A detection of the feature will constrain the water inventory of rocky exoplanets, providing important context in the assessment of their habitability.
\end{enumerate}
